Convolutional equivalent layer \\ for magnetic data processing

### 2022-05-12 manuscript/submission_review_final.tex
manuscrito final para fist look

%\documentclass[paper,twocolumn,twoside]{geophysics}
%\documentclass[manuscript,noblind]{geophysics}
\documentclass[manuscript]{geophysics}

@@

@@

%\ms{GEO-XXXX} % manuscript number

@@

We present a fast equivalent layer method for processing large-scale magnetic data. 
We demonstrate that the sensitivity matrix associated with an equivalent layer
of dipoles can be arranged to a Block-Toeplitz Toeplitz-Block (BTTB) structure for the 
case in which observations and dipoles are aligned on a 
horizontal and regularly-spaced grid.
The product of a BTTB matrix and an arbitrary vector represents a discrete 
convolution and can be efficiently computed via 2D Fast Fourier Transform.
In this case, the matrix-vector product uses only the elements forming the first column
of the BTTB matrix, saving computational time and memory. 
Our convolutional equivalent layer method uses this approach to compute 
the matrix-vector products in the iterative conjugate gradient algorithm with the purpose 
of estimating the physical-property distribution over the equivalent layer for 
large data sets.
Synthetic tests with a mid-size $100 \times 50$ grid of total-field anomaly data
show a decrease of $\approx 10^4$ in floating-point operations and $\approx 25\times$ 
in computation runtime of our method compared to the classical approach of solving
the least-squares normal equations via Cholesky decomposition. 
Faster results are obtained for millions of data, showing drastic decreases in computer memory usage
and runtime, allowing to perform magnetic data processing of large data sets on regular 
desktop computers. 
Our results also show that, compared to the classical Fourier approach, the magnetic
data processing with our method requires similar computation time, but produces significantly 
smaller border effects without using any padding scheme and also is more robust to 
deal with data on irregularly spaced points or on undulating observation surfaces.
A test with $1,310,000$ irregularly spaced field data over the Caraj{\'a}s Province, Brazil, 
confirms the efficiency of our method by taking $\approx 385.56$ seconds to estimate the physical-property
distribution over the equivalent layer and $\approx 2.64$ seconds to compute the upward-continuation.

@@

Large-scale data processing with tens of thousands of data is a reality in all areas of geophysics including the geophysical potential fields. 
The potential-field data processing includes convolution integrals which can be solved either in the space or Fourier domains.
The earliest techniques of potential-field data processing  were developed in the space domain.
For example, \cite{Peters1949} accomplished, in the space domain, the second and fourth derivatives of magnetic data and the upward- and downward-continuations of magnetic data by deriving coefficients that are
used in a  graphical convolution with the magnetic data.
However, the techniques for  processing potential-field data in space-domain were soon substituted by the Fourier-domain techniques. 
\cite{Dean1958} pointed out that the operations of second order derivatives, analytic continuation, smoothing, the removing of residuals or regionals and others for processing  potential-field data are similar to the electric filter circuits in Fourier domain.
\cite{Dean1958} was the first to develop the theory of linear filters in Fourier domain for gravity and magnetic processing and to present filters in Fourier domain \citep[][ see Table I, p 113]{Dean1958} 
for some theoretical geophysical operations (e.g., derivatives and upward- and downward-continuations).
\cite{GUNN1975} presented a comprehensive analysis of processing potential-field data in Fourier domain.

@@

Equations \ref{eq:BTTB_A_alpha_beta} and \ref{eq:Aq_block} define the general BTTB
structure of all matrix components $\mathbf{A_{\boldsymbol{\alpha\beta}}}$, but 
there are some differences between them.
Let us consider the matrix component $\mathbf{A}_{\boldsymbol{xx}}$, with elements
$a^{xx}_{ij}$ (equation \ref{eq:aij_alpha_beta}) defined by the second derivative calculated in
$h^{xx}_{ij}$ (equation \ref{eq:hxx_regular}). It can be easily verified from equations
\ref{eq:Delta_kij_symmetry} and \ref{eq:1_rij_symmetry} that $h^{xx}_{ij} = h^{xx}_{ji}$.
As a consequence, $a^{xx}_{ij} = a^{xx}_{ji}$, which means that 
\begin{equation}
	\mathbf{A}_{\boldsymbol{xx}} = \left( \mathbf{A}_{\boldsymbol{xx}} \right)^{\top}
	\label{eq:Axx_symmetry}
\end{equation}
for any grid orientation.
Now, let us investigate the elements $a^{xx}_{qp}$ forming the blocks $\mathbf{A}_{\boldsymbol{xx}}^{q}$.
For $x$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are defined 
by equations \ref{eq:q-x-oriented} and 
\ref{eq:p-x-oriented} and $a^{xx}_{qp}$ can be rewritten as follows:
\begin{equation}
	a^{xx}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
	\left( F_{x} u_{x} \right) \frac{3 \left( p \, \Delta_{x} \right)^{2}}{r_{qp}^{5}} - 
	\frac{1}{r_{qp}^{3}} \: ,
	\label{eq:aqp_xx_x_oriented}
\end{equation}
where
\begin{equation}
	\frac{1}{r_{qp}} = 
	\frac{1}{\sqrt{\left( p \, \Delta_{x} \right)^{2} + \left( q \, \Delta_{y} \right)^{2} + \Delta_{z}^{2}}} \: .
	\label{eq:1_rqp_x_oriented}
\end{equation}
For $y$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are 
defined by equations \ref{eq:q-y-oriented} and 
\ref{eq:p-y-oriented} and $a^{xx}_{qp}$ can be rewritten as follows:
\begin{equation}
	a^{xx}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
	\left( F_{x} u_{x} \right) \frac{3 \left( q \, \Delta_{x} \right)^{2}}{r_{qp}^{5}} - 
	\frac{1}{r_{qp}^{3}} \: ,
	\label{eq:aqp_xx_y_oriented}
\end{equation}
where
\begin{equation}
	\frac{1}{r_{qp}} = 
	\frac{1}{\sqrt{\left( q \, \Delta_{x} \right)^{2} + \left( p \, \Delta_{y} \right)^{2} + \Delta_{z}^{2}}} \: .
	\label{eq:1_rqp_y_oriented}
\end{equation}
From equations \ref{eq:aqp_xx_x_oriented}--\ref{eq:1_rqp_y_oriented}, we can easily verify that
\begin{equation}
	\mathbf{A}_{\boldsymbol{xx}}^{q} = \mathbf{A}_{\boldsymbol{xx}}^{(-q)}
	\label{eq:Axx_q_external_block_symmetry}
\end{equation}
and
\begin{equation}
	\mathbf{A}_{\boldsymbol{xx}}^{q} = \left(\mathbf{A}_{\boldsymbol{xx}}^{q} \right)^{\top} \: .
	\label{eq:Axx_q_internal_block_symmetry}
\end{equation}
Note that these symmetries are valid for 
any grid orientation.
From this results we conclude the matrix component 
$\mathbf{A}_{\boldsymbol{xx}}$ is \textit{symmetric-Block-Toeplitz symmetric-Toeplitz-Block} 
for any grid orientation.
The same reasoning can be used to show that matrices $\mathbf{A}_{\boldsymbol{yy}}$ and
$\mathbf{A}_{\boldsymbol{zz}}$ also have this symmetric structure.

@@

Let $\mathbf{A}_{\boldsymbol{xy}}$ be a matrix component with elements
$a^{xy}_{ij}$ (equation \ref{eq:aij_alpha_beta}) defined by the second derivative calculated in
$h^{xy}_{ij}$ (equation \ref{eq:hxy_regular}). It can be easily verified from equations
\ref{eq:Delta_kij_symmetry}--\ref{eq:1_rij_symmetry} that $h^{xy}_{ij} = h^{xy}_{ji}$.
As a consequence, $a^{xy}_{ij} = a^{xy}_{ji}$, which means that 
\begin{equation}
	\mathbf{A}_{\boldsymbol{xy}} = \left( \mathbf{A}_{\boldsymbol{xy}} \right)^{\top}
	\label{eq:Axy_symmetry}
\end{equation}
for any grid orientation.
For $x$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are defined 
by equations \ref{eq:q-x-oriented} and 
\ref{eq:p-x-oriented} and $a^{xy}_{qp}$ can be rewritten as follows:
\begin{equation}
	a^{xy}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
	\left( F_{x} u_{y} + F_{y} u_{x} \right) \frac{3 \left( p \, \Delta_{x} \right)\left( q \, \Delta_{y} \right)}{r_{qp}^{5}}
	\: ,
	\label{eq:aqp_xy_x_oriented}
\end{equation}
with $\tfrac{1}{r_{qp}}$ defined by equation \ref{eq:1_rqp_x_oriented}.
For $y$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are 
defined by equations \ref{eq:q-y-oriented} and 
\ref{eq:p-y-oriented} and $a^{xy}_{qp}$ can be rewritten as follows:
\begin{equation}
	a^{xy}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
	\left( F_{x} u_{y} + F_{y} u_{x} \right) \frac{3 \left( q \, \Delta_{x} \right)\left( p \, \Delta_{y} \right)}{r_{qp}^{5}} \: ,
	\label{eq:aqp_xy_y_oriented}
\end{equation}
with $\tfrac{1}{r_{qp}}$ defined by equation \ref{eq:1_rqp_y_oriented}.
From equations \ref{eq:1_rqp_x_oriented}, \ref{eq:1_rqp_y_oriented}, \ref{eq:aqp_xy_x_oriented} 
and \ref{eq:aqp_xy_y_oriented}, we can show that
\begin{equation}
	\mathbf{A}_{\boldsymbol{xy}}^{q} = -\mathbf{A}_{\boldsymbol{xy}}^{(-q)}
	\label{eq:Axy_q_external_block_symmetry}
\end{equation}
and 
\begin{equation}
	\mathbf{A}_{\boldsymbol{xy}}^{q} = -\left( \mathbf{A}_{\boldsymbol{xy}}^{q} \right)^{\top} \: .
	\label{eq:Axy_q_internal_block_symmetry}
\end{equation}
Note that these symmetries are valid for any grid orientation.
From this results we conclude the matrix component 
$\mathbf{A}_{\boldsymbol{xy}}$ is \textit{skew symmetric-Block-Toeplitz skew symmetric-Toeplitz-Block} 
for any grid orientation.

@@

Let $\mathbf{A}_{\boldsymbol{xz}}$ be a matrix component with elements
$a^{xz}_{ij}$ (equation \ref{eq:aij_alpha_beta}) defined by the second derivative calculated in
$h^{xz}_{ij}$ (equation \ref{eq:hxz_regular}). It can be easily verified from equations
\ref{eq:Delta_kij_symmetry}--\ref{eq:1_rij_symmetry} that $h^{xz}_{ij} = -h^{xz}_{ji}$.
As a consequence, $a^{xz}_{ij} = -a^{xz}_{ji}$, which means that 
\begin{equation}
	\mathbf{A}_{\boldsymbol{xz}} = -\left( \mathbf{A}_{\boldsymbol{xz}} \right)^{\top}
	\label{eq:Axz_symmetry}
\end{equation} 
for any grid orientation.
For $x$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are defined 
by equations \ref{eq:q-x-oriented} and 
\ref{eq:p-x-oriented} and $a^{xz}_{qp}$ can be rewritten as follows:
\begin{equation}
	a^{xz}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
	\left( F_{x} u_{z} + F_{z} u_{x} \right) \frac{3 \left( p \, \Delta_{x} \right) \Delta_{z}}{r_{qp}^{5}}
	\: ,
	\label{eq:aqp_xz_x_oriented}
\end{equation}
with $\tfrac{1}{r_{qp}}$ defined by equation \ref{eq:1_rqp_x_oriented}.
In this case, we can see that
\begin{equation}
	\mathbf{A}_{\boldsymbol{xz}}^{q} = \mathbf{A}_{\boldsymbol{xz}}^{(-q)}
	\label{eq:Axz_q_external_block_symmetry_x_oriented}
\end{equation}
and 
\begin{equation}
	\mathbf{A}_{\boldsymbol{xz}}^{q} = -\left( \mathbf{A}_{\boldsymbol{xz}}^{q} \right)^{\top} \: .
	\label{eq:Axz_q_internal_block_symmetry_x_oriented}
\end{equation}
This structure is called \textit{symmetric-Block-Toeplitz skew symmetric-Toeplitz-Block} and is 
valid only for $x$-oriented grids.
For $y$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are 
defined by equations \ref{eq:q-y-oriented} and 
\ref{eq:p-y-oriented} and $a^{xz}_{qp}$ can be rewritten as follows:
\begin{equation}
	a^{xz}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
	\left( F_{x} u_{z} + F_{z} u_{x} \right) \frac{3 \left( q \, \Delta_{x} \right) \Delta_{z}}{r_{qp}^{5}} \: ,
	\label{eq:aqp_xz_y_oriented}
\end{equation}
with $\tfrac{1}{r_{qp}}$ defined by equation \ref{eq:1_rqp_y_oriented}.
Now, we conclude that
\begin{equation}
	\mathbf{A}_{\boldsymbol{xz}}^{q} = -\mathbf{A}_{\boldsymbol{xz}}^{(-q)}
	\label{eq:Axz_q_external_block_symmetry_y_oriented}
\end{equation}
and 
\begin{equation}
	\mathbf{A}_{\boldsymbol{xz}}^{q} = \left( \mathbf{A}_{\boldsymbol{xz}}^{q} \right)^{\top} \: .
	\label{eq:Axz_q_internal_block_symmetry_y_oriented}
\end{equation}
This structure is called \textit{skew symmetric-Block-Toeplitz symmetric-Toeplitz-Block} and is 
valid only for $y$-oriented grids.

@@

Figure \ref{fig:flops} shows a comparison between 
$f_{classical}$ (equation \ref{eq:flops-classical-method}), 
$f_{cgls}$ (equation \ref{eq:flops-standard-cgls}) and 
$f_{conv}$ (equation \ref{eq:flops-convolutional-method})
for different numbers of observation points up to $1,000,000$. As we can see, 
the total flops count associated with our method is $7$ orders of magnitude smaller 
than that associated with the classical method and $3$ orders of magnitude smaller than
that associated with the standard CGLS method when using a maximum number
of iterations $N^{it} = 50$.

@@

Our convolutional equivalent layer method requires a regular data grid located on a 
horizontal and flat observation surface.
Here, we evaluate the performance of our method by applying it to simulated airborne magnetic 
surveys formed by
i) a regular data grid on a flat surface;
ii) irregular data grids on a flat surface; and 
iii) regular data grid on undulating surfaces.
Note that the simulated surveys in (ii) and (iii) violate the premises of our method.

@@

%% regular grid on undulating surfaces

Figures \ref{fig:synthetic_data_comparison_v2}d and \ref{fig:synthetic_data_comparison_v2}e 
show the simulated surveys on the same regular grid as shown in Figure 
\ref{fig:synthetic_data_comparison_v2}a (upper panel). The difference is that observation points
are located no longer on a flat, but on undulating surfaces.
For the survey shown in Figure \ref{fig:synthetic_data_comparison_v2}d, the $z$ coordinates 
of the undulating surface are defined by a sequence of pseudorandom Gaussian noise having mean 
$-900$ m and standard deviation equal to $5\%$ of $900$ m, which corresponds to $45$ m.
For the survey shown in Figure \ref{fig:synthetic_data_comparison_v2}e, the standard deviation 
is equal to $10\%$ of $900$ m, which corresponds to $90$ m.
The noise-corrupted total-field anomalies of these simulated surveys (second panels in Figures 
\ref{fig:synthetic_data_comparison_v2}d and \ref{fig:synthetic_data_comparison_v2}e) are calculated 
on their corresponding undulating surfaces (upper panels in Figures 
\ref{fig:synthetic_data_comparison_v2}d and \ref{fig:synthetic_data_comparison_v2}e),
on the same regular grid shown in Figure \ref{fig:synthetic_data_comparison_v2}a,
with pseudorandom Gaussian noise added having null mean and standard deviation of $0.2961$ nT.

@@

Figure \ref{fig:synthetic_residuals_convergence_comparison_v2} show the 
difference between the simulated (second row in Figure \ref{fig:synthetic_data_comparison_v2})
and predicted data (not shown) obtained by using the classical (the upper row) and 
our method (the second row). From now on, we designate this difference as data residuals. 
The lower row in Figure \ref{fig:synthetic_residuals_convergence_comparison_v2} shows the 
convergence curve of our method.

>>>>

%\documentclass[paper,twocolumn,twoside]{geophysics}
\documentclass[manuscript,noblind]{geophysics}
%\documentclass[manuscript]{geophysics}
%\documentclass[manuscript,revised]{geophysics}

@@

\title{Convolutional equivalent layer \\ for magnetic data processing}

@@

\ms{GEO-2021-0599} % manuscript number

\address{
\footnotemark[1]Observatório Nacional, \\
77 General José Cristino St, \\
Rio de Janeiro, RJ, 20921-400 \\
}
\author{Diego Takahashi\footnotemark[1], Vanderlei C. Oliveira Jr.\footnotemark[1] and Valéria C. F. Barbosa\footnotemark[1]}

%\footer{Example}
\lefthead{Takahashi et al.}
\righthead{Magnetic convolutional equivalent layer}

@@

We present a fast equivalent layer method for processing large-scale magnetic data. 
We demonstrate that the sensitivity matrix associated with an equivalent layer
of dipoles can be arranged to a Block-Toeplitz Toeplitz-Block (BTTB) structure for the 
case where observations and dipoles are aligned on a 
horizontal and regularly-spaced grid.
The product of a BTTB matrix and an arbitrary vector represents a discrete 
convolution and can be efficiently computed via 2D Fast Fourier Transform.
In this case, the matrix-vector product uses only the elements forming the first column
of the BTTB matrix, saving computational time and memory. 
Our convolutional equivalent layer method uses this approach to compute 
the matrix-vector products in the iterative conjugate gradient algorithm with the purpose 
of estimating the physical-property distribution over the equivalent layer for 
large data sets.
Synthetic tests of total-field anomaly data
show a decrease in both floating-point operations and computation runtime 
of our method compared to the classical approach of solving
the least-squares normal equations via Cholesky decomposition.
Faster results are obtained for millions of data, showing drastic decreases in computer memory usage
and runtime, allowing to perform magnetic data processing of large data sets on regular 
desktop computers. 
Our results also show that, compared to the classical Fourier approach, the magnetic
data processing with our method requires similar computation time, but produces significantly 
smaller border effects without using any padding scheme and is also more robust to 
deal with data on irregularly spaced points or on irregularly observation surfaces.
A test with irregularly spaced field data over the Caraj{\'a}s Province, Brazil, 
confirms the efficiency of our method by estimating the physical-property
distribution over the equivalent layer and computing the upward-continuation.

@@

The potential-field data processing includes convolution integrals which can be solved either in the space or Fourier domains.
The earliest techniques of potential-field data processing  were developed in the space domain.
For example, \cite{Peters1949} accomplished, in the space domain, the second and fourth derivatives of magnetic data and the upward- and downward-continuations of magnetic data by deriving coefficients that are
used in a  graphical convolution with the magnetic data.
However, the techniques for  processing potential-field data in space-domain were soon substituted by the Fourier-domain techniques. 
\cite{Dean1958} pointed out that the operations of second order derivatives, analytic continuation, smoothing, the removing of residuals or regionals and others for processing  potential-field data are similar to the electric filter circuits in Fourier domain.
\cite{Dean1958} was the first to develop the theory of linear filters in Fourier domain for gravity and magnetic processing and to present filters in Fourier domain \citep[][ see Table I, p 113]{Dean1958} 
for some theoretical geophysical operations (e.g., derivatives and upward- and downward-continuations).
\cite{GUNN1975} presented a comprehensive analysis of processing potential-field data in the Fourier domain.

@@

Equations \ref{eq:BTTB_A_alpha_beta} and \ref{eq:Aq_block} define the general BTTB
structure of all matrix components $\mathbf{A_{\boldsymbol{\alpha\beta}}}$, but 
there are some differences between them.
Let us consider the matrix component $\mathbf{A}_{\boldsymbol{xx}}$, with elements
$a^{xx}_{ij}$ (equation \ref{eq:aij_alpha_beta}) defined by the second derivative calculated in
$h^{xx}_{ij}$ (equation \ref{eq:hxx_regular}). It can be easily verified from equations
\ref{eq:Delta_kij_symmetry} and \ref{eq:1_rij_symmetry} that $h^{xx}_{ij} = h^{xx}_{ji}$.
Consequently, $a^{xx}_{ij} = a^{xx}_{ji}$, which means that 
\begin{equation}
	\mathbf{A}_{\boldsymbol{xx}} = \left( \mathbf{A}_{\boldsymbol{xx}} \right)^{\top}
	\label{eq:Axx_symmetry}
\end{equation}
for any grid orientation.
Now, let us investigate the elements $a^{xx}_{qp}$ forming the blocks $\mathbf{A}_{\boldsymbol{xx}}^{q}$.
For $x$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are defined 
by equations \ref{eq:q-x-oriented} and 
\ref{eq:p-x-oriented} and $a^{xx}_{qp}$ can be rewritten as follows:
\begin{equation}
	a^{xx}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
	\left( F_{x} u_{x} \right) \frac{3 \left( p \, \Delta_{x} \right)^{2}}{r_{qp}^{5}} - 
	\frac{1}{r_{qp}^{3}} \: ,
	\label{eq:aqp_xx_x_oriented}
\end{equation}
where
\begin{equation}
	\frac{1}{r_{qp}} = 
	\frac{1}{\sqrt{\left( p \, \Delta_{x} \right)^{2} + \left( q \, \Delta_{y} \right)^{2} + \Delta_{z}^{2}}} \: .
	\label{eq:1_rqp_x_oriented}
\end{equation}
For $y$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are 
defined by equations \ref{eq:q-y-oriented} and 
\ref{eq:p-y-oriented} and $a^{xx}_{qp}$ can be rewritten as follows:
\begin{equation}
	a^{xx}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
	\left( F_{x} u_{x} \right) \frac{3 \left( q \, \Delta_{x} \right)^{2}}{r_{qp}^{5}} - 
	\frac{1}{r_{qp}^{3}} \: ,
	\label{eq:aqp_xx_y_oriented}
\end{equation}
where
\begin{equation}
	\frac{1}{r_{qp}} = 
	\frac{1}{\sqrt{\left( q \, \Delta_{x} \right)^{2} + \left( p \, \Delta_{y} \right)^{2} + \Delta_{z}^{2}}} \: .
	\label{eq:1_rqp_y_oriented}
\end{equation}
From equations \ref{eq:aqp_xx_x_oriented}--\ref{eq:1_rqp_y_oriented}, we can easily verify that
\begin{equation}
	\mathbf{A}_{\boldsymbol{xx}}^{q} = \mathbf{A}_{\boldsymbol{xx}}^{(-q)}
	\label{eq:Axx_q_external_block_symmetry}
\end{equation}
and
\begin{equation}
	\mathbf{A}_{\boldsymbol{xx}}^{q} = \left(\mathbf{A}_{\boldsymbol{xx}}^{q} \right)^{\top} \: .
	\label{eq:Axx_q_internal_block_symmetry}
\end{equation}
Note that these symmetries are valid for 
any grid orientation.
From this results we conclude the matrix component 
$\mathbf{A}_{\boldsymbol{xx}}$ is \textit{symmetric-Block-Toeplitz symmetric-Toeplitz-Block} 
for any grid orientation.
The same reasoning can be used to show that matrices $\mathbf{A}_{\boldsymbol{yy}}$ and
$\mathbf{A}_{\boldsymbol{zz}}$ also have this symmetric structure.

@@

Let $\mathbf{A}_{\boldsymbol{xy}}$ be a matrix component with elements
$a^{xy}_{ij}$ (equation \ref{eq:aij_alpha_beta}) defined by the second derivative calculated in
$h^{xy}_{ij}$ (equation \ref{eq:hxy_regular}). It can be easily verified from equations
\ref{eq:Delta_kij_symmetry}--\ref{eq:1_rij_symmetry} that $h^{xy}_{ij} = h^{xy}_{ji}$.
Then, $a^{xy}_{ij} = a^{xy}_{ji}$, which means that 
\begin{equation}
	\mathbf{A}_{\boldsymbol{xy}} = \left( \mathbf{A}_{\boldsymbol{xy}} \right)^{\top}
	\label{eq:Axy_symmetry}
\end{equation}
for any grid orientation.
For $x$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are defined 
by equations \ref{eq:q-x-oriented} and 
\ref{eq:p-x-oriented} and $a^{xy}_{qp}$ can be rewritten as follows:
\begin{equation}
	a^{xy}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
	\left( F_{x} u_{y} + F_{y} u_{x} \right) \frac{3 \left( p \, \Delta_{x} \right)\left( q \, \Delta_{y} \right)}{r_{qp}^{5}}
	\: ,
	\label{eq:aqp_xy_x_oriented}
\end{equation}
with $\tfrac{1}{r_{qp}}$ defined by equation \ref{eq:1_rqp_x_oriented}.
For $y$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are 
defined by equations \ref{eq:q-y-oriented} and 
\ref{eq:p-y-oriented} and $a^{xy}_{qp}$ can be rewritten as follows:
\begin{equation}
	a^{xy}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
	\left( F_{x} u_{y} + F_{y} u_{x} \right) \frac{3 \left( q \, \Delta_{x} \right)\left( p \, \Delta_{y} \right)}{r_{qp}^{5}} \: ,
	\label{eq:aqp_xy_y_oriented}
\end{equation}
with $\tfrac{1}{r_{qp}}$ defined by equation \ref{eq:1_rqp_y_oriented}.
From equations \ref{eq:1_rqp_x_oriented}, \ref{eq:1_rqp_y_oriented}, \ref{eq:aqp_xy_x_oriented} 
and \ref{eq:aqp_xy_y_oriented}, we can show that
\begin{equation}
	\mathbf{A}_{\boldsymbol{xy}}^{q} = -\mathbf{A}_{\boldsymbol{xy}}^{(-q)}
	\label{eq:Axy_q_external_block_symmetry}
\end{equation}
and 
\begin{equation}
	\mathbf{A}_{\boldsymbol{xy}}^{q} = -\left( \mathbf{A}_{\boldsymbol{xy}}^{q} \right)^{\top} \: .
	\label{eq:Axy_q_internal_block_symmetry}
\end{equation}
Note that these symmetries are valid for any grid orientation.
From this results we conclude the matrix component 
$\mathbf{A}_{\boldsymbol{xy}}$ is \textit{skew symmetric-Block-Toeplitz skew symmetric-Toeplitz-Block} 
for any grid orientation.

@@

Let $\mathbf{A}_{\boldsymbol{xz}}$ be a matrix component with elements
$a^{xz}_{ij}$ (equation \ref{eq:aij_alpha_beta}) defined by the second derivative calculated in
$h^{xz}_{ij}$ (equation \ref{eq:hxz_regular}). It can be easily verified from equations
\ref{eq:Delta_kij_symmetry}--\ref{eq:1_rij_symmetry} that $h^{xz}_{ij} = -h^{xz}_{ji}$.
Therefore, $a^{xz}_{ij} = -a^{xz}_{ji}$, which means that 
\begin{equation}
	\mathbf{A}_{\boldsymbol{xz}} = -\left( \mathbf{A}_{\boldsymbol{xz}} \right)^{\top}
	\label{eq:Axz_symmetry}
\end{equation} 
for any grid orientation.
For $x$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are defined 
by equations \ref{eq:q-x-oriented} and 
\ref{eq:p-x-oriented} and $a^{xz}_{qp}$ can be rewritten as follows:
\begin{equation}
	a^{xz}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
	\left( F_{x} u_{z} + F_{z} u_{x} \right) \frac{3 \left( p \, \Delta_{x} \right) \Delta_{z}}{r_{qp}^{5}}
	\: ,
	\label{eq:aqp_xz_x_oriented}
\end{equation}
with $\tfrac{1}{r_{qp}}$ defined by equation \ref{eq:1_rqp_x_oriented}.
In this case, we can see that
\begin{equation}
	\mathbf{A}_{\boldsymbol{xz}}^{q} = \mathbf{A}_{\boldsymbol{xz}}^{(-q)}
	\label{eq:Axz_q_external_block_symmetry_x_oriented}
\end{equation}
and 
\begin{equation}
	\mathbf{A}_{\boldsymbol{xz}}^{q} = -\left( \mathbf{A}_{\boldsymbol{xz}}^{q} \right)^{\top} \: .
	\label{eq:Axz_q_internal_block_symmetry_x_oriented}
\end{equation}
This structure is called \textit{symmetric-Block-Toeplitz skew symmetric-Toeplitz-Block} and is 
valid only for $x$-oriented grids.
For $y$-oriented grids (Figure \ref{fig:regular-grids}), the block indices $q$ and $p$ are 
defined by equations \ref{eq:q-y-oriented} and 
\ref{eq:p-y-oriented} and $a^{xz}_{qp}$ can be rewritten as follows:
\begin{equation}
	a^{xz}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
	\left( F_{x} u_{z} + F_{z} u_{x} \right) \frac{3 \left( q \, \Delta_{x} \right) \Delta_{z}}{r_{qp}^{5}} \: ,
	\label{eq:aqp_xz_y_oriented}
\end{equation}
with $\tfrac{1}{r_{qp}}$ defined by equation \ref{eq:1_rqp_y_oriented}.
Now, we conclude that
\begin{equation}
	\mathbf{A}_{\boldsymbol{xz}}^{q} = -\mathbf{A}_{\boldsymbol{xz}}^{(-q)}
	\label{eq:Axz_q_external_block_symmetry_y_oriented}
\end{equation}
and 
\begin{equation}
	\mathbf{A}_{\boldsymbol{xz}}^{q} = \left( \mathbf{A}_{\boldsymbol{xz}}^{q} \right)^{\top} \: .
	\label{eq:Axz_q_internal_block_symmetry_y_oriented}
\end{equation}
This structure is called \textit{skew symmetric-Block-Toeplitz symmetric-Toeplitz-Block} and is 
valid only for $y$-oriented grids.

@@

Figure \ref{fig:flops} shows a comparison between 
$f_{classical}$ (equation \ref{eq:flops-classical-method}), 
$f_{cgls}$ (equation \ref{eq:flops-standard-cgls}) and 
$f_{conv}$ (equation \ref{eq:flops-convolutional-method})
for different numbers of observation points up to $1,000,000$. As we can see, 
the total flops count associated with our method is seven orders of magnitude smaller 
than that associated with the classical method and three orders of magnitude smaller than
that associated with the standard CGLS method when using a maximum number
of iterations $N^{it} = 50$.

@@

Our convolutional equivalent layer method requires a regular data grid located on a 
horizontal and flat observation surface.
Here, we evaluate the performance of our method by applying it to simulated airborne magnetic 
surveys formed by
i) a regular data grid on a flat surface;
ii) irregular data grids on a flat surface; and 
iii) regular data grids on $x$- and $y$-directions but, irregular on the $z$-direction (irregular surface).
Note that the simulated surveys in (ii) and (iii) violate the premises of our method, implying 
that matrix $\mathbf{A}$ is not BTTB.

@@

%% regular grid on irregular surfaces

Figures \ref{fig:synthetic_data_comparison_v2}d and \ref{fig:synthetic_data_comparison_v2}e 
show the simulated surveys on the same regular grid as shown in Figure 
\ref{fig:synthetic_data_comparison_v2}a (upper panel). The difference is that observation points
are located no longer on a flat surface, but irregular in the $z$-direction.
For the survey shown in Figure \ref{fig:synthetic_data_comparison_v2}d, the $z$ coordinates 
of the irregular surface are defined by a sequence of pseudorandom Gaussian noise having mean 
$-900$ m and standard deviation equal to $5\%$ of $900$ m, which corresponds to $45$ m.
For the survey shown in Figure \ref{fig:synthetic_data_comparison_v2}e, the standard deviation 
is equal to $10\%$ of $900$ m, which corresponds to $90$ m.
The noise-corrupted total-field anomalies of these simulated surveys (second panels in Figures 
\ref{fig:synthetic_data_comparison_v2}d and \ref{fig:synthetic_data_comparison_v2}e) are calculated 
on their corresponding irregular surfaces (upper panels in Figures 
\ref{fig:synthetic_data_comparison_v2}d and \ref{fig:synthetic_data_comparison_v2}e),
on the same regular grid shown in Figure \ref{fig:synthetic_data_comparison_v2}a,
with pseudorandom Gaussian noise added having null mean and standard deviation of $0.2961$ nT.

@@

Figure \ref{fig:synthetic_residuals_convergence_comparison_v2} show the 
difference between the simulated (second row in Figure \ref{fig:synthetic_data_comparison_v2})
and predicted data (not shown) obtained with the classical (the upper row) and 
our method (the second row). From now on, we designate this difference as data residuals. 
The lower row in Figure \ref{fig:synthetic_residuals_convergence_comparison_v2} shows the 
convergence curve of our method.

The data residuals using the classical method (equation \ref{eq:classical-method})  
are shown in the upper panel of Figure \ref{fig:synthetic_residuals_convergence_comparison_v2}a, 
with mean $0.4118$ nT and standard deviation $0.3780$ nT. This process took $17.10$ seconds.
Using our method, the data residuals (the middle panel in Figure
\ref{fig:synthetic_residuals_convergence_comparison_v2}a) have mean $0.9972$ nT and standard
deviation $1.3904$ nT. In this case, however, the processing time was only $0.25$ seconds.
As expected, the Euclidean norm of the data residuals produced by our method 
(lower panel in Figure \ref{fig:synthetic_residuals_convergence_comparison_v2}a) decreases. 
The convergence criterion was satisfied at iteration $51$.
%This result shows that, in practice, our method converges way before $N$ iterations,
%where $N$ is the number of observations.
%Setting the convergence to $N$ iterations, besides being unnecessary, it also demands a larger 
%computer processing time, even in this synthetic test with a small number 
%of $N = 5,\, 000$ observations.
